Supplemental Online Materials:\\\textit{Meta-Analysis of the `Ironic' Effects of Intergroup Contact}

### 2021-05-14 manuscript/manuscript.tex
submit

\noindent Critics have argued that intergroup contact, psychology's
most-researched paradigm for reducing prejudice, has the `ironic' effect
of stifling support for social change in disadvantaged groups. We
conducted a preregistered meta-analytic test of this effect across 97
studies with 138 samples of 212,189 disadvantaged-group members. As
predicted, intergroup contact was, on average, associated with less
perceived injustice (\(r = -.07\)), collective action (\(r = -.06\)),
and support for reparative policies (\(r = -.07\)) in disadvantaged
groups. However, these associations were small, variable, and consistent
with alterative explanations. Moderator analyses found that these
associations were strongest in studies on short-term migration and
post-colonial relations, that measured intergroup contact directly, and
that involved adults. After controlling for the positive association of
negative contact with support for social change, positive contact was
not (negatively) associated with any outcome in studies that measured
both predictors. We close by highlighting open questions about the
relation between intergroup contact and social change.

@@

It is ironic that psychology's most-researched paradigm for reducing
prejudice---intergroup contact---should hinder, rather than help, social
change. Hundreds of studies have confirmed that intergroup contact is
associated with less negative feelings toward outgroup members (for a
meta-analysis, see Pettigrew \& Tropp, 2006). Most of these studies have
focused on advantaged-group members' contact experiences with
disadvantaged-group members (Tropp \& Pettigrew, 2005), guided by the
idea that reducing advantaged-group members' prejudice will prevent
interpersonal discrimination. Reducing prejudice, however, is rarely
enough to overcome institutional and structural discrimination. Instead,
as the 2020 Black Lives Matter protests have shown, social change often
requires political mobilization by the disadvantaged. Critics have
argued that, by fostering harmony and reducing conflict, intergroup
contact might distract disadvantaged-group members from the injustice
they face and undermine their motivation for social change (Dixon et
al., 2005; Reicher, 2007; Wright, 2003). By stifling disadvantaged-group
members' opposition to injustice, intergroup contact might thus have the
`ironic' (Saguy et al., 2009), `sedative' (Çakal et al., 2011), or
`paradoxical' (Dixon et al., 2010) effect of hindering social change.

Supporting this argument, initial studies provided evidence that, among
disadvantaged-group members, intergroup contact reduces perceptions of
injustice (Saguy et al., 2009), discourages collective action (Çakal et
al., 2011), and diminishes support for reparative policies (Dixon et
al., 2007). Others, however, have pointed to conflicting evidence (Poore
et al., 2002) and argued that intergroup contact might instead render
discrimination more salient and, thereby, increase disadvantaged-group
members' support for social change (Pettigrew et al., 2011). A decade
after the initial studies, many more studies have examined the
relationships between intergroup contact and support for social change
in disadvantaged groups. We present a systematic review and
meta-analysis of this growing literature. Specifically, we evaluate the
evidence for three hypotheses about how intergroup contact might reduce
support for social change.

Perceived injustice---that is, perceiving unjust group-based deprivation
or discrimination against the disadvantaged ingroup---is a prerequisite
to instigating social change (van Zomeren et al., 2008). Researchers
have argued that positive contact with advantaged-group members could
contradict perceptions of personal discrimination, undermine negative
characterizations of the advantaged outgroup, and, thereby, make group
discrimination appear less plausible (Dixon et al., 2010). Positive
contact could further reduce perceived injustice by emphasizing
commonalities over differences (Saguy et al., 2009) and by motivating
disadvantaged-group members to adopt system-justifying ideologies
(Sengupta \& Sibley, 2013). To the extent that it reduces perceived
injustice, intergroup contact hinders social change.

Collective action---that is, any action directed at improving the
conditions of the disadvantaged ingroup (Wright et al., 1990)---is
crucial to achieving social change because advantaged-group members
rarely surrender their advantages without sustained pressure from
disadvantaged-group members (Blumer, 1958; for a review, see Dixon et
al., 2012). Researchers have argued that intergroup contact could
discourage disadvantaged-group members from engaging in collective
action by diminishing their perceived discrimination (Tropp et al.,
2012), by blurring boundaries between the disadvantaged ingroup and the
advantaged outgroup (Saguy et al., 2009), by reducing negative attitudes
toward the advantaged outgroup (Wright \& Lubensky, 2009), and by
quelling anger about inequality (Tausch et al., 2015) or toward the
advantaged outgroup (Hayward et al., 2018). To the extent that it
discourages collective action, intergroup contact hinders social change.

Reparative policies (e.g., affirmative action) can be effective means to
reduce social inequality. Disadvantaged-group members, who stand to
benefit from their implementation, are important advocates for these
policies. Researchers have argued that intergroup contact could diminish
disadvantaged-group members' support for redistributive policies by
creating sympathy for advantaged-group members (Dixon et al., 2007), by
fostering positive outgroup attitudes and reducing attention to
inequality (Saguy et al., 2009), and by increasing their endorsement of
system-justifying ideologies (Sengupta \& Sibley, 2013). To the extent
that it diminishes support for redistributive policies, intergroup
contact hinders social change.

We conducted a preregistered meta-analysis of the evidence for the
hypothesized negative effects of intergroup contact on perceived
injustice, collective action, and policy support in disadvantaged
groups. Our first objective was to systematically review the available
evidence and to evaluate its strengths and limitations. Our second
objective was to synthesize the available evidence to establish the
direction and magnitude of the relationships between intergroup contact
and the three outcomes. Doing so allowed us, first, to determine whether
the available evidence supports a negative or positive (see Pettigrew et
al., 2011) relationship between contact and support for social change
and, second, to estimate the strength of this relationship. Our third
objective was to explore the variability of this relationship across
study settings, designs, and other potential moderators. For example, we
examined whether any effects were confined to Western, Educated,
Industrialized, Rich, Democratic (WEIRD) settings, the cultural context
of most psychological research (Henrich et al., 2010). Our fourth
objective was to assess how robust the available evidence was to
publication bias.

In addition, we collated the evidence for two alternative accounts of
this relationship. First, we considered studies that measured both
ingroup and outgroup contact as predictors of the three outcomes.
Sengupta et al. (2015) argued that contact with other members of the
disadvantaged ingroup would increase political mobilization in
disadvantaged groups. As outgroup contact places limits on how much
ingroup contact disadvantaged-group members can have (Pfister et al.,
2020), the mobilizing effect of ingroup contact is an alternative
explanation for negative associations between outgroup contact and
support for social change. Second, we considered studies that measured
both negative and positive (outgroup) contact as predictors of the three
outcomes. Researchers (Hayward et al., 2018; Reimer et al., 2017) found
negative contact with advantaged-group members to be associated with
greater perceived discrimination and stronger collective action
intentions. Reimer et al. (2017) argued that, to the extent
disadvantaged-group members who have more positive contact also tend to
have less negative contact, the mobilizing effect of negative contact is
an alternative explanation for negative associations between outgroup
contact and support for social change.

@@

We included studies if they measured or manipulated the quantity of,
quality of, and opportunity for contact with members of outgroups that
are relatively advantaged compared to the participants' ingroup. We
considered opportunity for contact, as it is a potential precursor to
and proxy for face-to-face contact, but not imagined or indirect
contact. Similarly, we included studies that measured intergroup contact
indirectly by, for example, asking what proportion of someone's friends
were \emph{not} from the participants' ingroup.

@@

\textbf{\emph{Perceived Injustice.}} We included studies that measured
perceptions that one is discriminated against because of one's group
membership, that one's group faces discrimination in society, that one's
group is relatively deprived compared to other groups, or that the
deprivation and/or discrimination faced by one's group is unjust and
illegitimate. We also considered expectations of fair treatment as the
reverse of perceived injustice. We included studies that measured
personal and/or group discrimination or a mixture of both.

@@

Bayesian inference involves choosing a likelihood function and prior
distributions. A likelihood function links the observed data to one or
more model parameters and states how likely the observed data are given
different values of said model parameters. Prior distributions state how
plausible different values of said model parameters are before
considering the observed data. Bayesian inference applies Bayes' theorem
to update prior distributions in light of the observed data to produce
posterior distributions. Other than \emph{p}-values and confidence
intervals, the resulting posterior distributions have a straightforward
interpretation as stating how plausible different values of the model
parameters are given the observed data. We report point estimates, based
on the median of posterior samples, and uncertainty intervals, based on
the quantiles of posterior samples, that enclose the 95\% most plausible
estimates. In addition, we report the posterior probability, based on
the proportion of posterior samples below zero, that an estimated mean
effect size is negative.

@@

Figure 1 shows a flow diagram illustrating our search strategy, study
selection, and data collection. Our preregistered search strategy
returned 2,610 unique records from electronic databases. Of these, we
excluded 2,379 (91\%) ineligible records after screening titles,
abstracts, and keywords. Of the remaining 231 records, we excluded 116
(50\%) ineligible records after reviewing full-text manuscripts. We
supplemented these records with 15 unpublished studies solicited from
researchers and 6 records that cited at least three relevant works. Of
136 eligible studies, we had to exclude 38 (28\%) studies for which we
could not extract or impute any relevant effect size. At each stage, we
also excluded studies that used the same data as another study. Our
final sample comprised effect sizes from \(J = 97\) studies spanning
\(N = 212,189\) participants in \(I = 138\) samples.

Figure 2 provides a qualitative overview of the literature. Even though
most participants came from studies conducted in India (\(N = 104,639\))
and the United States (\(N = 60,983\); Figure 2a), most samples were
collected in North America (\(I = 48\)) and Europe (\(I = 41\)), with
few samples from South America and Africa (Figure 2b). Most samples
focused on relative inequalities that resulted from long-term migration
(\(I = 57\)), slavery (\(I = 19\)), colonization (\(I = 19\)), and
short-term migration (\(I = 14\)). With few exceptions, studies used
observational, cross-sectional survey designs. Samples sizes ranged from
64 to 49,764 (\(\textit{Mdn} = 283\)). Unlike most psychological
research, 2 in 5 samples used probability or representative sampling. Of
all samples, 54\% (\(I = 75\)) were collected with the intention to
examine the effects of intergroup contact on the outcomes considered in
this meta-analysis.\footnote{Many studies examined acculturation
  processes (Berry, 1997) and measured both intergroup contact and
  perceived injustice without the relationship between the two being of
  interest.} Together, these observations highlight both limitations and
strengths of the empirical literature.

@@

\textbf{\emph{Perceived Injustice.}} Across 202,741 participants from
124 samples in 85 studies, we found strong evidence for a weak
association (\(r = -.07, [-.10, -.04]\)) between intergroup contact and
perceived injustice, with \(>99.9\%\) of posterior samples for the mean
correlation coefficient falling below zero. We found evidence that
correlation coefficients varied across studies
(\(\tau_J = .14, [.12, .17]\)) and across samples within studies
(\(\tau_I = .08, [.05, .12]\)). Based on these analyses, we predicted
that 80\% of studies would result in correlation coefficients between
-.25 and .10 and that researchers would need sample sizes of at least
2,549 participants to find significant associations (\(\alpha = .05\),
two-sided) in 80\% of their studies.\footnote{Sample sizes are based on
  posterior predictions from the three models, which implied that, for
  80\% of studies, the absolute correlation coefficient would be
  \(|r| > .039\) for perceived injustice, \(|r| > .045\) for collective
  action, and \(|r| > .035\) for policy support.}

@@

There is an emerging consensus that intergroup contact has the `ironic'
effect of undermining support for social change in disadvantaged groups.
We conducted a preregistered meta-analytic test of this effect across 96
studies with 137 samples of 211,360 disadvantaged-group members. We
found that, based on the available evidence, the associations of
intergroup contact with perceived injustice (\(\Pr (r < 0) > 99.9\%\)),
collective action (\(\Pr (r < 0) = 93.6\%\)), and support for reparative
policies (\(\Pr (r < 0) = 98.1\%\)) were, on average, much more likely
to be negative than positive. Thus, our meta-analysis seems to support
the emerging consensus. We use the rest of this section to argue why
this conclusion might be premature.

@@

A combination of moderators explained about a third of the between-study
variance in the association between intergroup contact and perceived
injustice. We found that, on average, this association was negative only
in studies of adults that measured contact \emph{directly} by asking,
for example, about the number of friends from the advantaged outgroup.
This suggests that, as hypothesized, the `ironic' effects result from
direct contact and not from mere exposure or other contextual factors.
We also found that, on average, effect sizes were greater for studies in
post-colonial settings or on short-term migration than for studies in
other contexts. Future research should systematically investigate
variance in how contact affects support for social change across
settings and cultures.

Third, the available evidence is almost entirely from cross-sectional,
observational studies and thus consistent with alternative explanations
for the observed associations. Support for social change could reduce
intergroup contact rather than the other way around. For example,
disadvantaged-group members involved in collective action might
purposefully avoid forming friendships with advantaged-group members.
Alternatively, the association between intergroup contact and support
for social change could be spurious with both being caused by an
unobserved confounder. For example, a disadvantaged-group member's
socioeconomic status might both expose them to more advantaged-group
members and reduce their perception of injustice. Future research should
prioritize longitudinal studies to confirm the direction of the observed
associations and (field) experiments to rule out confounding and other
alternative explanations.\footnote{Published research includes only
  three longitudinal studies (Koschate et al., 2012; Reimer et al.,
  2017; Tropp et al., 2012), all of which conflated within-person change
  and between-person stability (Hamaker et al., 2015), two experimental
  studies with a no-contact control condition (Becker et al., 2013;
  Droogendyk et al., 2016), and two intervention studies (Reimer et al.,
  2021; Shani \& Boehnke, 2017).}

Further, the observed association between intergroup contact and support
for social change could be confounded by an alternative predictor that
\emph{increases} support for social change. Supporting Reimer et al.'s
(2017) argument, we found that positive intergroup contact was not
associated with support for social change after controlling for negative
contact which, in turn, was associated with \emph{greater} support for
social change. We found mixed evidence for ingroup contact as an
alternative explanation for the `ironic' effects of intergroup contact
(Sengupta et al., 2015) but note that few studies measured both ingroup
and outgroup contact. Going forward, researchers should clarify which
aspects of intergroup contact should theoretically affect support for
social change---and include measures that allow testing both the
hypothesized and competing explanations.

Fourth, we cannot rule out that biases in the literature caused our
meta-analysis to overestimate the `ironic' effects of intergroup
contact. Publication bias could have prevented studies that found
positive or non-significant associations from entering the published
literature. We used state-of-the-art methods to detect and correct for
publication bias, yet did not find conclusive evidence for or against
publication bias. That said, 61\% of samples were either not collected
with the intention to study the effects of intergroup contact on a
relevant outcome (45\%) or from unpublished studies (32\%)---and thus
unlikely to have been affected by publication bias. Another bias that
could have affected our findings is that researchers made data-dependent
decisions when analyzing data that inflate false-positive findings
(Gelman \& Loken, 2014). Future research should use preregistration and
other methods to prevent undisclosed flexibility in data collection and
analysis.\footnote{To date, one preregistered study has been published
  (Hässler et al., 2020).}

>>>>

\noindent Growing evidence suggests that intergroup contact, psychology's
most-researched paradigm for reducing prejudice, might have the `ironic'
effect of reducing support for social change in disadvantaged groups. We
conducted a preregistered meta-analysis of this effect across 97 studies
with 138 samples of 212,189 disadvantaged-group members. As predicted,
intergroup contact was, on average, associated with less perceived
injustice (\(r = -.07\)), collective action (\(r = -.06\)), and support
for reparative policies (\(r = -.07\)). Moderator analyses found that
associations were strongest in studies of adults that measured
intergroup contact directly and examined short-term migration or
post-colonial intergroup relations. After controlling for the positive
association of negative contact with support for social change, positive
contact was not associated with any outcome in studies that measured
both predictors. We close by highlighting open questions about the
relationship between intergroup contact and social change.

@@

It is ironic that psychology's most-researched paradigm for reducing
prejudice---intergroup contact---should hinder, rather than help, social
change. Hundreds of studies have confirmed that intergroup contact is
associated with less negative feelings toward outgroup members (for a
meta-analysis, see Pettigrew \& Tropp, 2006). Most studies have focused
on advantaged-group members' contact experiences with
disadvantaged-group members (Tropp \& Pettigrew, 2005), guided by the
idea that reducing advantaged-group members' prejudice will prevent
interpersonal discrimination. Reducing prejudice, however, is rarely
enough to overcome institutional and structural discrimination. Instead,
as the 2020 Black Lives Matter protests have shown, social change often
requires political mobilization by the disadvantaged. Critics have
argued that, by fostering harmony and reducing conflict, intergroup
contact might distract disadvantaged-group members from the injustice
they face and undermine their motivation for social change (Dixon et
al., 2005; Reicher, 2007; Wright, 2003). By reducing disadvantaged-group
members' opposition to injustice, intergroup contact might thus have the
`ironic' (Saguy et al., 2009), `sedative' (Çakal et al., 2011), or
`paradoxical' (Dixon et al., 2010) effect of hindering social change.

Supporting this argument, initial studies provided evidence that contact
with advantaged-group members reduces disadvantaged-group members'
perceptions of injustice (Saguy et al., 2009), collective action (Çakal
et al., 2011), and support for reparative policies (Dixon et al., 2007).
Others, however, have pointed to conflicting evidence (Poore et al.,
2002) and argued that intergroup contact might instead make
discrimination more salient and, thereby, increase disadvantaged-group
members' support for social change (Pettigrew et al., 2011). In the
decade since the initial studies, many more studies have examined the
relationships between intergroup contact and support for social change
in disadvantaged groups. We present a systematic review and
meta-analysis of this growing literature. Specifically, we evaluate the
evidence for three hypotheses about how intergroup contact might reduce
support for social change.

Perceived injustice---that is, perceiving unjust group-based deprivation
or discrimination against a disadvantaged ingroup---is a prerequisite to
instigating social change (van Zomeren et al., 2008). Researchers have
argued that positive contact with advantaged groups could contradict
perceptions of personal discrimination and, thereby, make group
discrimination appear less plausible (Dixon et al., 2010). Positive
contact could further reduce perceived injustice by emphasizing
commonalities over differences (Saguy et al., 2009) and by motivating
disadvantaged-group members to adopt system-justifying ideologies
(Sengupta \& Sibley, 2013). To the extent that it reduces perceived
injustice, intergroup contact reduces support for social change.

Collective action---that is, any action directed at improving the
conditions of the disadvantaged ingroup (Wright et al., 1990)---is
crucial to achieving social change because advantaged-group members
rarely surrender their advantages without sustained pressure from
disadvantaged-group members (Blumer, 1958; for a review, see Dixon et
al., 2012). Researchers have argued that intergroup contact could
discourage disadvantaged-group members from engaging in collective
action by diminishing their perceived discrimination (Tropp et al.,
2012), by blurring boundaries between the disadvantaged ingroup and the
advantaged outgroup (Saguy et al., 2009), by reducing negative attitudes
toward the advantaged outgroup (Wright \& Lubensky, 2009), and by
quelling their anger (Hayward et al., 2018; Tausch et al., 2015). To the
extent that it discourages collective action, intergroup contact reduces
support for social change.

Reparative policies (e.g., affirmative action) can be effective means to
reduce social inequality. Disadvantaged-group members, who stand to
benefit from their implementation, are important advocates for these
policies. Researchers have argued that intergroup contact could diminish
disadvantaged-group members' support for redistributive policies by
creating sympathy for advantaged-group members (Dixon et al., 2007), by
fostering positive outgroup attitudes and reducing attention to
inequality (Saguy et al., 2009), and by increasing their endorsement of
system-justifying ideologies (Sengupta \& Sibley, 2013). To the extent
that it diminishes support for redistributive policies, intergroup
contact reduces support for social change.

Our first objective was to systematically review the available evidence
and to evaluate its strengths and limitations. Our second objective was
to synthesize the available evidence to establish the direction and
magnitude of the relationships between intergroup contact and the three
outcomes. Doing so allowed us to determine whether the available
evidence supports a negative or positive (see Pettigrew et al., 2011)
relationship between contact and support for social change and to
estimate the strength of this relationship. Our third objective was to
explore the variability of this relationship across study settings,
designs, and other potential moderators. For example, we examined
whether any effects were confined to Western, Educated, Industrialized,
Rich, Democratic (WEIRD) countries, the cultural context of most
psychological research (Henrich et al., 2010). Our fourth objective was
to assess how robust the available evidence was to publication bias.

In addition, we examined the evidence for two alternative accounts of
this relationship. First, we considered studies that measured both
ingroup and outgroup contact as predictors of the three outcomes.
Sengupta et al. (2015) argued that contact with other members of the
disadvantaged ingroup would increase political mobilization. As outgroup
contact places limits on how much ingroup contact disadvantaged-group
members can have (Pfister et al., 2020), the mobilizing effect of
ingroup contact is an alternative explanation for negative associations
between outgroup contact and support for social change.

Second, we considered studies that measured both negative and positive
(outgroup) contact as predictors of the three outcomes. Researchers
(Hayward et al., 2018; Reimer et al., 2017) found negative contact with
advantaged-group members to be associated with greater perceived
discrimination and stronger collective action intentions. Reimer et al.
(2017) argued that, to the extent disadvantaged-group members who have
more positive contact also tend to have less negative contact, the
mobilizing effect of negative contact is an alternative explanation for
negative associations between outgroup contact and support for social
change.

@@

We included studies if they measured or manipulated the quantity of,
quality of, and opportunity for contact with members of outgroups that
are relatively advantaged compared to the participants' ingroup. We
considered opportunity for contact, as it is a potential precursor to
and proxy for face-to-face contact, but not imagined or extended
contact. Similarly, we included studies that measured intergroup contact
indirectly by, for example, asking what proportion of someone's friends
were \emph{not} from the participants' ingroup.

@@

\textbf{\emph{Perceived Injustice.}} We included studies that measured
perceptions that one is discriminated against because of one's group
membership, that one's group faces discrimination in society, that one's
group is relatively deprived compared to other groups, or that the
deprivation and/or discrimination faced by one's group is unjust and
illegitimate. We also considered expectations of fair treatment as the
reverse of perceived injustice. We included studies that measured
personal or group discrimination or a mixture of both.

@@

Bayesian inference involves choosing a likelihood function and prior
distributions. A likelihood function links the observed data to one or
more model parameters and states how likely the observed data are given
different values of said model parameters. Prior distributions state how
plausible different values of said model parameters are before
considering the observed data. Bayesian inference applies Bayes' theorem
to update prior distributions in light of the observed data to produce
posterior distributions. Other than \emph{p}-values and confidence
intervals, the resulting posterior distributions have a straightforward
interpretation as stating how plausible different values of the model
parameters are given the observed data. We report point estimates, based
on the median of posterior samples, and uncertainty intervals, based on
the quantiles of posterior samples, that enclose the 95\% most plausible
estimates. In addition, we report the posterior probability, based on
the proportion of posterior samples below zero, that the mean effect
size is negative.

@@

Figure 1 shows a flow diagram illustrating our search strategy, study
selection, and data collection. Our preregistered search strategy
returned 2,610 unique records from electronic databases. Of these, we
excluded 2,379 (91\%) ineligible records after screening titles,
abstracts, and keywords. Of the remaining 231 records, we excluded 116
(50\%) ineligible records after reviewing full-text manuscripts. We
supplemented these records with 14 unpublished studies and 6 records
that cited at least three relevant works. Of 135 eligible studies, we
had to exclude 38 (28\%) studies for which we could not extract or
impute any relevant effect size. At each stage, we also excluded studies
that used the same data as another study. Our final sample comprised
effect sizes from \(J = 97\) studies spanning \(N = 212,189\)
participants in \(I = 138\) samples (for a complete list of studies, see
SOM-U).

Figure 2 provides a qualitative overview of the literature. Even though
most participants came from studies conducted in India (\(N = 104,639\))
and the United States (\(N = 60,983\); Figure 2a), most samples were
collected in North America (\(I = 48\)) and Europe (\(I = 41\)), with
few samples from South America and Africa (Figure 2b). Most samples
focused on relative inequalities that resulted from long-term migration
(\(I = 57\)), slavery (\(I = 19\)), colonization (\(I = 19\)), and
short-term migration (\(I = 14\)). With few exceptions, studies used
observational, cross-sectional survey designs. Samples sizes ranged from
64 to 49,764 (\(\textit{Mdn} = 283\)). Unlike most psychological
research, 2 in 5 samples used probability or representative sampling. Of
all samples, 54\% (\(I = 75\)) were collected with the intention to
examine the effects of intergroup contact on the outcomes considered in
this meta-analysis. Many studies instead examined acculturation
processes (Berry, 1997) and measured intergroup contact and perceived
injustice without hypothesizing that the former would influence the
latter. Together, these observations highlight both limitations and
strengths of the empirical literature.

@@

\textbf{\emph{Perceived Injustice.}} Across 202,741 participants from
124 samples in 85 studies, we found strong evidence for a weak
association (\(r = -.07, [-.10, -.04]\)) between intergroup contact and
perceived injustice, with \(>99.9\%\) of posterior samples for the mean
correlation coefficient falling below zero. We found evidence that
correlation coefficients varied across studies
(\(\tau_J = .14, [.12, .17]\)) and across samples within studies
(\(\tau_I = .08, [.05, .12]\)). Based on these analyses, we predicted
that 80\% of studies would result in correlation coefficients between
-.25 and .10 and that researchers would need sample sizes of at least
2,549 participants to find significant associations (\(\alpha = .05\),
two-sided) in 80\% of their studies.\footnote{Sample sizes are based on
  posterior predictions which implied that, for 80\% of studies, the
  absolute correlation coefficient would be \(|r| > .039\) for perceived
  injustice, \(|r| > .045\) for collective action, and \(|r| > .035\)
  for policy support.}

@@

There is an emerging consensus that intergroup contact has the `ironic'
effect of undermining support for social change in disadvantaged groups.
We conducted a preregistered meta-analytic test of this effect across 96
studies with 137 samples of 211,360 disadvantaged-group members. We
found that, based on the available evidence, the associations of
intergroup contact with perceived injustice (\(\Pr (r < 0) > 99.9\%\)),
collective action (\(\Pr (r < 0) = 93.6\%\)), and support for reparative
policies (\(\Pr (r < 0) = 98.1\%\)) were, on average, much more likely
to be negative than positive. Thus, our meta-analysis seems to support
the emerging consensus. However, our findings qualify this conclusion in
several important ways.

@@

A combination of moderators explained about a third of the between-study
variance in the association between intergroup contact and perceived
injustice. We found that, on average, this association was negative only
in studies of adults that measured contact \emph{directly} (e.g., by
asking about the number of friends from the advantaged outgroup). This
suggests that, as hypothesized, the `ironic' effects result from direct
contact and not from other contextual factors. We also found that, on
average, effect sizes were greater for studies in post-colonial settings
or on short-term migration than for studies in other contexts. Future
research should systematically investigate variance in how contact
affects support for social change across settings and cultures.

Third, the available evidence is almost entirely from cross-sectional,
observational studies and thus consistent with alternative explanations
for the observed associations. Support for social change could reduce
intergroup contact rather than the other way around. For example,
disadvantaged-group members involved in collective action might avoid
forming friendships with advantaged-group members. Alternatively, the
observed associations could be spurious with both being caused by an
unobserved confounder. For example, a disadvantaged-group member's
socioeconomic status might both expose them to more advantaged-group
members and reduce their perception of injustice. Future research should
prioritize longitudinal studies to confirm the direction of the observed
associations and (field) experiments to rule out confounding and other
alternative explanations.\footnote{Published research includes only
  three longitudinal studies (Koschate et al., 2012; Reimer et al.,
  2017; Tropp et al., 2012), all of which conflated within-person change
  and between-person stability (Hamaker et al., 2015), two experimental
  studies with a no-contact control condition (Becker et al., 2013;
  Droogendyk et al., 2016), and two intervention studies (Reimer et al.,
  2021; Shani \& Boehnke, 2017).}

Fourth, the observed association between intergroup contact and support
for social change could be confounded by an alternative predictor that
\emph{increases} support for social change. Supporting Reimer et al.'s
(2017) argument, we found that positive intergroup contact was not
associated with support for social change after controlling for negative
contact which, in turn, was associated with \emph{greater} support for
social change. We found mixed evidence for ingroup contact as an
alternative explanation for the `ironic' effects of intergroup contact
(Sengupta et al., 2015) but note that few studies measured both ingroup
and outgroup contact. Going forward, researchers should clarify which
aspects of intergroup contact should theoretically affect support for
social change---and include measures that allow testing both the
hypothesized and competing explanations.

Finally, we cannot rule out that biases in the literature caused our
meta-analysis to overestimate the `ironic' effects of intergroup
contact. Publication bias could have prevented studies that found
positive or non-significant associations from entering the published
literature. We used various methods to detect and correct for
publication bias, yet did not find conclusive evidence for or against
publication bias. That said, 61\% of samples were either not collected
with the intention to study the effects of intergroup contact on a
relevant outcome (45\%) or from unpublished studies (32\%)---and thus
unlikely to have been affected by publication bias. Another bias that
could have affected our findings is that researchers made data-dependent
decisions when analyzing data that inflate false-positive findings
(Gelman \& Loken, 2014). Future research should use preregistration and
other methods to prevent undisclosed flexibility in data collection and
analysis.\footnote{Only one preregistered study has been published
  (Hässler et al., 2020).}

### 2021-05-14 manuscript/supplemental-materials.Rmd
submit

First, we ran a series of Bayesian random-effects meta-regression models to estimate how much of the between-samples heterogeneity was explained by specific categorical moderator variables (`run_moderator_analyses.R`). Figures S1 and S2 show the results for collective action and policy support. Comparisons were inconclusive because we had an insufficient number of effect sizes per category. Second, we used meta-regression trees to discover interactions between moderator variables that best explained heterogeneity in effect sizes [@li_meta-cart_2017; @li_multiple_2020]. As recommended, we ran random-effects meta-regression tree analyses using the look-ahead strategy [@li_multiple_2020]. We set the pruning parameter to $c = 0$ because, for this analysis, we prioritized exploration over error control (`run_exploratory_moderator_analyses.R`). For both collective action and policy support, the algorithm found that no moderator or interaction of moderators explained between-samples heterogeneity. This result is not surprising, however, as both analyses used fewer than 40 effect sizes, the minimum number required for the algorithm to perform well in detecting even simple interaction effects [@li_meta-cart_2017]. As both moderator analyses were inconclusive for collective action and policy support, we only report results for perceived injustice in the main text.

>>>>

First, we ran a series of Bayesian random-effects meta-regression models to estimate how much of the between-samples heterogeneity was explained by specific categorical moderator variables (`run_moderator_analyses.R`). Figures S1 and S2 show the results for collective action and policy support. Comparisons were inconclusive because we had an insufficient number of effect sizes per category. We ran another three random-effects meta-regression models with Muthukrishna et al.'s [-@muthukrishna_beyond_2020] measure of cultural distance, where available, as a continuous moderator variable. Because cultural distance was a country-level moderator variable, we used two-level random-effects models in which we estimated country-specific deviations from the mean effect size as well as sample-specific deviations from the country-specific effect size. Figure 5 (in the main text) show the inconclusive results for all three outcomes.

Second, we used meta-regression trees to discover interactions between moderator variables that best explained heterogeneity in effect sizes [@li_meta-cart_2017; @li_multiple_2020]. As recommended, we ran random-effects meta-regression tree analyses using the look-ahead strategy [@li_multiple_2020]. We set the pruning parameter to $c = 0$ because, for this analysis, we prioritized exploration over error control (`run_exploratory_moderator_analyses.R`). For both collective action and policy support, the algorithm found that no moderator or interaction of moderators explained between-samples heterogeneity. This result is not surprising, however, as both analyses used fewer than 40 effect sizes, the minimum number required for the algorithm to perform well in detecting even simple interaction effects [@li_meta-cart_2017]. As all moderator analyses were inconclusive for collective action and policy support, we only report results for perceived injustice in the main text.

### 2021-05-14 manuscript/supplemental-materials.tex
submit

First, we ran a series of Bayesian random-effects meta-regression models
to estimate how much of the between-samples heterogeneity was explained
by specific categorical moderator variables
(\texttt{run\_moderator\_analyses.R}). Figures S1 and S2 show the
results for collective action and policy support. Comparisons were
inconclusive because we had an insufficient number of effect sizes per
category. Second, we used meta-regression trees to discover interactions
between moderator variables that best explained heterogeneity in effect
sizes (Li et al., 2017, 2020). As recommended, we ran random-effects
meta-regression tree analyses using the look-ahead strategy (Li et al.,
2020). We set the pruning parameter to \(c = 0\) because, for this
analysis, we prioritized exploration over error control
(\texttt{run\_exploratory\_moderator\_analyses.R}). For both collective
action and policy support, the algorithm found that no moderator or
interaction of moderators explained between-samples heterogeneity. This
result is not surprising, however, as both analyses used fewer than 40
effect sizes, the minimum number required for the algorithm to perform
well in detecting even simple interaction effects (Li et al., 2017). As
both moderator analyses were inconclusive for collective action and
policy support, we only report results for perceived injustice in the
main text.

@@

>>>>

First, we ran a series of Bayesian random-effects meta-regression models
to estimate how much of the between-samples heterogeneity was explained
by specific categorical moderator variables
(\texttt{run\_moderator\_analyses.R}). Figures S1 and S2 show the
results for collective action and policy support. Comparisons were
inconclusive because we had an insufficient number of effect sizes per
category. We ran another three random-effects meta-regression models
with Muthukrishna et al.'s (2020) measure of cultural distance, where
available, as a continuous moderator variable. Because cultural distance
was a country-level moderator variable, we used two-level random-effects
models in which we estimated country-specific deviations from the mean
effect size as well as sample-specific deviations from the
country-specific effect size. Figure 5 (in the main text) show the
inconclusive results for all three outcomes.

Second, we used meta-regression trees to discover interactions between
moderator variables that best explained heterogeneity in effect sizes
(Li et al., 2017, 2020). As recommended, we ran random-effects
meta-regression tree analyses using the look-ahead strategy (Li et al.,
2020). We set the pruning parameter to \(c = 0\) because, for this
analysis, we prioritized exploration over error control
(\texttt{run\_exploratory\_moderator\_analyses.R}). For both collective
action and policy support, the algorithm found that no moderator or
interaction of moderators explained between-samples heterogeneity. This
result is not surprising, however, as both analyses used fewer than 40
effect sizes, the minimum number required for the algorithm to perform
well in detecting even simple interaction effects (Li et al., 2017). As
all moderator analyses were inconclusive for collective action and
policy support, we only report results for perceived injustice in the
main text.

@@

\leavevmode\hypertarget{ref-muthukrishna_beyond_2020}{}%
Muthukrishna, M., Bell, A. V., Henrich, J., Curtin, C. M., Gedranovich,
A., McInerney, J., \& Thue, B. (2020). Beyond western, educated,
industrial, rich, and democratic (WEIRD) psychology: Measuring and
mapping scales of cultural and psychological distance.
\emph{Psychological Science}, \emph{31}(6), 678--701.
\url{https://doi.org/10.1177/0956797620916782}
